\begin{thebibliography}{30}
	\bibitem{towards}
	\emph{S. Aguzzoli and P. Codara}, \textbf{Towards an Algebraic Topos Semantics for Three-valued Gödel Logic} 2021 IEEE International Conference on Fuzzy Systems (FUZZ-IEEE), Luxembourg, 2021, pp. 1-6.
	
	\bibitem{fuzzy}
	\emph{Cintula P., Hájek P., Noguera C.} (2011).\textbf{ Handbook of Mathematical Fuzzy Logic} - volume 2. London : College Publications.
	
	\bibitem{metamath}
	\emph{Hájek, Petr} (1998). \textbf{Metamathematics of Fuzzy Logic}. Dordrecht, Boston and London: Kluwer Academic Publishers.
	
	\bibitem{firstorder}
	\emph{Baaz, Matthias et al.} \textbf{First-order Gödel logics.} Ann. Pure Appl. Log. 147 (2006): 23-47.
	
	\bibitem{manyval}
	\emph{S.Aguzzoli, P. Codara \& V.Marra.} \textbf{Gödel-Dummett logic, the category of forests and topoi} 
	ManyVal 2019 Department of C.S. University of Bucharest.
	
	\bibitem{recursive}
	\emph{S. Aguzzoli and P. Codara}, \textbf{Recursive formulas to compute coproducts of finite Gödel algebras and related structures} 2016 IEEE International Conference on Fuzzy Systems (FUZZ-IEEE), Vancouver, BC, Canada, 2016, pp. 201-208.
	
	\bibitem{computing}
	\emph{D’Antona, Ottavio M. \& Marra, Vincenzo} (2006). \textbf{Computing coproducts of finitely presented Gödel algebras}. Annals of Pure and Applied Logic 142 (1):202-211.
	
	\bibitem{stone}
	\emph{Stone, M.H.} \textbf{The theory of representation of Boolean algebras.} Trans. Amer. Math Soc. 40, 1936.
	
	\bibitem{goldblatt}
	\emph{Goldblatt, R. I.} (1982). \textbf{Topoi: The Categorial Analysis of Logic}. British Journal for the Philosophy of Science 33 (1):95-97.
	
	\bibitem{awodey}
	\emph{Awodey, Steve} (2006). \textbf{Category Theory}. Oxford, England: Oxford University Press.
	
	\bibitem{godel}
	\emph{Solomon Feferman, John W. Dawson, Stephen C. Kleene, Gregory H. Moore, Robert M. Solovay, and Jean van Heijenoort} (Eds.). 1986. \textbf{Kurt Godel: collected works. Vol. 1: Publications 1929-1936}. Oxford University Press, Inc., USA.
	
	
	\bibitem{lambekscott}
	\emph{Lambek, J. \& Scott, P. J.} (1989). \textbf{Introduction to Higher Order Categorical Logic}. Journal of Symbolic Logic.
	
	\bibitem{borceaux}
	\emph{Borceux, F.} (1994). \textbf{Handbook of Categorical Algebra 3: Categories of Sheaves} (Encyclopedia of Mathematics and its Applications). Cambridge: Cambridge University Press.
	
	\bibitem{lisboa}
	\emph{Pedro Filipe.} \textbf{A topoi characterization of Gödel intermediate logics.} Master’s
	thesis, Instituto Superior Técnico, 2017
	
	\bibitem{elephant}
	\emph{Johnstone, Peter T.} (2002). \textbf{Sketches of an Elephant: A Topos Theory Compendium}, Volume 1. Oxford, England: Clarendon Press.
	
	\bibitem{heyting}
	\emph{A. Heyting} (1930), \textbf{Die Formalen Regeln der intuitionistischen Logik} Sitzungsberichte der Preussischen Akademie von Wissenschaften. Physikalisch Mathematische Klasse 42–56. A.
	
	\bibitem{maclane}
	\emph{MacLane, S.} (1971). \textbf{Categories for the Working Mathematician}. New York: Springer-Verlag.
	
	\bibitem{mtl}
	\emph{Esteva, Francesc \& Godo, Lluis.} (2001) \textbf{Monoidal t-norm Based Logic: Towards a Logic for Left-continuous t-norms}. Fuzzy Sets and Systems. Published by Elsevier BV. 
	
	
\end{thebibliography}


